\loesung{
\begin{enumerate}[a)]
\item 
Reading the P(good credit risk) panel (left): All categories show a high average probability of being classified as low risk (good). However, \texttt{male : divorced/separated} has the lowest average probability ($\approx$0.68 vs.\ $\approx$0.72 for other groups), suggesting the model treats this group slightly differently.
\item
ALE plots are preferable to PDPs when features are correlated, because PDPs average predictions over the entire marginal distribution, including potentially unrealistic feature combinations that rarely or never occur in the data (the extrapolation problem). ALE avoids this by only using predictions within each feature's local neighborhood and computing local differences rather than absolute predictions. This makes ALE unbiased even with correlated features. Additionally, ALE is computationally more efficient, requiring $O(2n)$ model evaluations vs.\ $O(n \cdot g)$ for PDP (where $g$ is the grid size).
% SAY in CLASS:
% Since \texttt{personal\_status\_sex} is a categorical feature, we cannot 
% use the Pearson correlation coefficients (even for 
% numerical features it is not always a good idea (see in-class exercise)). 
% For two categorical features we can use a $\Chi^2$-test 
% and for categorical vs. a numerical features we could rely on a one-way ANOVA F-test. 
\item 
The following pseudocode computes the pairwise sum of the absolute differences of relative 
frequencies in the categories of a categorical feature $x_j$ based on a 
feature $x_k$

%\lstinputlisting{"rsrc/get_diff_cat().R"}

\begin{algorithm}[ht]
\caption{\texttt{get\_diff\_cat()}}
\begin{algorithmic}[1]
\Require \texttt{feature.k}: values of categorical feature for which relative frequencies per class are calculated
\Require \texttt{feature.j}: values of categorical feature for which similarity based on \texttt{feature.k} should be assessed
\State \texttt{dists} $\gets$ unique class combinations of \texttt{feature.j}
\State \texttt{x.count} $\gets$ number of observations per class of \texttt{feature.j}
\State \texttt{A} $\gets$ relative cross table of \texttt{feature.j} and \texttt{feature.k} weighted by \texttt{x.count} (per class of \texttt{feature.j} relative frequencies of classes of \texttt{feature.k} should sum up to 1)
\State \texttt{dist} $\gets$ sum up distances of probability distributions per unique class combination specified in \texttt{dists}
\State \textbf{return} \texttt{dist}, \texttt{dists}
\end{algorithmic}
\end{algorithm}

For our task at hand, we obtain the following distances
% \lstinputlisting{"rsrc/run_get_diff_cat.R"}
\lstinputlisting{"rsrc/run_get_diff_cat.txt"}

Overall, the following ordering of \texttt{personal\_status\_sex} was returned by the method: 
% \lstinputlisting{"rsrc/run_order_levels.R"}
\lstinputlisting{"rsrc/run_order_levels.txt"}


The ordering seems to be feasible, since categories including females are close to each other 
and also categories with males. 
Also the ordering of males according to their relationship status seems to make sense, since typically 
the process is: single, then married and then divorced :-).

\item \textbf{Bonus:} ALE and PDP are global interpretation tools which base their insights on averages (of predictions or 
prediction differences) over whole test sets.
Indeed vulnerable groups are typically not the majority of a population but have a low proportion, 
and biases might be overlooked.
Therefore, local explanation tools should be consulted, in addition to these methods in order to 
identify pointwise biases or discriminatory behavior.

\end{enumerate}
}