\aufgabe{LIME Quiz}{
\begin{enumerate}
	    \item Which of the following statement(s) about local surrogate models is/are correct?  
	\begin{enumerate}
		\item
		Surrogate models produced by LIME should have the same prediction as the model to be explained for the entire training dataset.
		\item
		The choice of the sampling process and the definition of locality are important hyperparameters of LIME that have a large impact on the behavior of the method.
		\item
		LIME does not require any adaptions to be applicable to deep learning models for image data.
		\item
		LIME requires the surrogate model to use all available features - a selection of features is not allowed.
	\end{enumerate}
	
	\item Which of the following statement(s) about the weighting scheme in LIME is/are correct?
	\begin{enumerate}
		\item
		Observations closer to the instance of interest receive higher weights.
		\item
		If the kernel width for the exponential kernel is set to infinity, all observations receive a proximity measure/weight of $1$ independent of their distance to $\xv$.
		\item
		All perturbed samples are weighted equally regardless of distance.
		\item
		The kernel width controls how quickly weights decay with distance.
	\end{enumerate}

	
	\item Which of the following statement(s) about the limitations of LIME is/are correct?
	\begin{enumerate}
		\item
		LIME explanations can be unstable for small changes in the input.
		\item
		LIME guarantees globally faithful explanations of the model.
		\item
		Correlated features require uniformly sampling from univariate distributions.
		\item 
		If many observations lie in the neighborhood of the point of interest, these observations should be used as training points for the surrogate model. 
	\end{enumerate}

	
%\item Which of the following properties describe a \textbf{good superpixel segmentation} for LIME explanations?
%\begin{enumerate}
%	\item
%	A superpixel should align with meaningful object boundaries in the image.
%	\item
%	A superpixel should be randomly assigned to pixels to avoid bias.
%	\item
%	A superpixel should be internally homogeneous (similar color/texture).
%	\item
%	A superpixel should always cover the entire image with equal square blocks.
%\end{enumerate}
%	
%	
%\item How does the size of superpixels influence LIME image explanations?
%\begin{enumerate}
%	\item
%	Explanations become more fine-grained and detailed if the size of superpixels increases.
%	\item
%	Explanations become more coarse and less precise if the size of superpixels increases.
%	\item
%	It becomes harder to localize specific influential regions if the size of superpixels decreases.
%	\item
%	Increasing size of the superpixels decreases the number of features in the surrogate model.
%\end{enumerate}

\end{enumerate}
}
