\textbf{Quiz:}
\begin{enumerate}
%        \item The value of the PDP at a point $x_j$, corresponds to the point-wise average of the values of the ICE curves at this point. 
%        \item The PDP of a feature provides information about possible interaction effects of the feature. 
%        \item ICE curves of a feature for multiple data points provide information about possible interaction effects of the feature with others. 
%        \item If we center the ICE/PDPs for categorical features, the expected changes always refer to a selected reference category. 
%        \item ALE plots are based on conditional distributions, PDPs on marginal distributions.
%        \item If features are uncorrelated, ALE plots are equal to PDPs. 
%        \item ALE plots are faster to compute than PDPs.
        
    \item What is the problem of a PDP when interactions between features are present? What about extrapolation?
    \item How do PDPs and ICE curves correspond to each other?
    \item What problem do we need to keep in mind when using centered ICE/PDP for categorical features? 
    \item M-Plots handle correlated data well and do not suffer from extrapolation. What disadvantage does this method have?
    \item Name the advantages and disadvantages of the different marginal effect (ME) methods (dME, fME and AME) in comparison to the other feature effect methods.
    \item Name the advantages of ALE over PDP.
    \item Can you think of a situation in which ALE equals PDP?
    \item How does the interpretation between M-Plots and ALE differ?
    \item How are marginal effects different than all the other feature effect methods (ICE, PDP, M-plots, ALE)?


    \item
    You fitted a model that should predict the value of a property depending on the number of rooms and square meters.
    You want to compute feature effects using the following methods: PDP, M-plots and ALE plots. 
    Which of the following strategies reflect which method? \\
    The feature effect for a 30 m$^2$ flat corresponds to... 
    \begin{enumerate}[a)]
        \item ... what the model predicts on average for flats that also have around 30 m$^2$, e.g., 28 m$^2$ to 32 m$^2$. 
        \item ... how the model prediction changes on average when flats with 28 m$^2$ to 32 m$^2$ have 32 m$^2$ vs. 28 m$^2$. 
        \item ... what the model predicts on average if all properties in the dataset have 30 m$^2$. 
\end{enumerate}

\end{enumerate}