\loesung{

\begin{enumerate}

    \item 
    Both PDP and ALE plots show a strong linear effect of $x_1$, where higher values of $x_1$ lead to higher values of predicted value.
    The PDP and ALE plots of $x_2$ show a strong decreasing effect of $x_2$ on the prediction. 
    The PDP of $x_2$ shows a steep jump for large values of $x_2$, while the ALE plot shows a strong linear effect over the whole range of values of $x_2$.
    Interpretation at the point $x_1 = 0.5$:
    \begin{itemize}
        \item PDP: the model predicts on average a value of around $2.3$ for $y$ if for all data instances $x_1 = 0.5$.
        \item ALE: the model predicts on average an increase of around $2.5$ of $y$ for data instances with $x_1 = 0.5$ 
        compared to the average prediction.
    \end{itemize}
    
    \item 
    PDPs assume that features are uncorrelated.
    % We know from the GAM output above - as well as the scatter plot - that $x_1$ and $x_2$ are highly correlated.
    But we know from the form of the DGP (or we could find out using a scatter plot, or by trying to model the relationship between $x_1$ and $x_2$, e.g. with a GAM) that $x_1$ and $x_2$ are highly correlated.
    Since PDPs extrapolate over predictions of artificial points that are out of distribution, the interpretation might be misleading, especially in areas with low data density (values of $x_2$ smaller than 0, or for high values of $x_2$ and small for $x_1$), or if the model contains interactions.
    ALE plots, on the other hand, do not predict in regions that are far away from the training data and, therefore, do not suffer from the extrapolation issue of PDPs. 
    
    \item[\textbf{Bonus:}]
    The observed distribution of each feature is visualized at the bottom of the plot.
    For $x_1$, it shows a homogeneous distribution between $-1$ and 1, as expected.
    The empirical distribution of $x_2$ centers in the interval $[0,1]$, which we also expected.
    But from the DGP given we know that $-1 \leq x_1 \leq 1$ always holds and therefore any values of $x_2$ outside of the interval $[0,1]$ are solely caused by the noise term $\delta$.
    
    Now, if $\delta$ is normally distributed with standard variance $0.04$, then $|\delta|>0.1$ is extremely unlikely, hence $x_2 < -0.1$ and $x_2 > 1.1$ should only occur in a very small number of cases.
    In contrast, we observe that inside the intervals $[-1.5, -1]$ and $[1, 1.5]$, the density of $x_2$ is quite high.
    This strongly indicates that the data and the plots were actually generated from a different DGP than the one given.
    
\end{enumerate}
  
}
