\textbf{Quiz:}
\begin{enumerate}

    \item 
    % With respect to what property of a function or model do functional decompositions decompose it?
    Functional decompositions are decompositions with respect to what property of a function?
    % interaction structure
    \item 
    What are functional decompositions useful for?
    \item 
    Why is the general definition of functional decompositions insufficient for practical use? How can these problems be alleviated?
    % Alternative: Why are additional constraints besides the general definition necessary to arrive at a useful functional decomposition?
    % See standard fANOVA: Remove lower-order terms from higher-order terms, and try to include as much as possible / reasonable in the lower-order terms
    \item 
    True or false: When given a function explicitly through a formula and applying a functional decomposition algorithm on it, one always recovers the original components of the formula as components in the resulting decomposition.
    \item 
    How do PDPs and general multidimensional PD-functions relate to the components in classical fANOVA?
    \item 
    In what order are the components in the classical fANOVA computed?
    \item 
    How are Sobol indices defined, and what property of the classical fANOVA decomposition does one need for them?
    \item
    You have trained a random forest classifier. Can you easily and directly compute the functional decomposition of the classifier and how would you do it? Or why can you not?
    % Trivial / strightforward decomposition of trees also works for tree ensembles like e.g. a random forest
    % Problem: This naive decomposition is not very insightful and usually does not capture main effects => Still need one of the model-agnostic methods to compute a functional decomposition
    \item 
    How is the interaction type of a node inside a decision tree defined?
    \item 
    Why are problems of computing and using the classical fANOVA in practice?
    % computationally expensive (as for any functional decomposition), independence assumption, high computational burden of PD-functions (can be alleviated via calculating all PD-functions in an hierarchical manner from the highest-order to the lowest-order, using dynamical programming
    % \item
    % Why are problems of computing and using functional decompositions in practice, and how can one deal with them?
    % Too many features => expensive & complicated to understand => sparse decomposition or only focus on specific interactions of interest, or models like EBMs
    % Only for tabular data, extracting tabular data from raw data is difficult to address
    % Need to choose one method, all of which have individual disadvantages
    % \item 
    % What do the vanishing condition and the orthogonality condition intuitively mean?
    \item 
    Which other IML methods discussed in this lecture so far have used or are related to the concept of functional decomposition?
    % LMs, GLMs, GAMs, EBMs, RPFs, PDPs, ALE, Shapley
    \item 
    Given a model $\fh$, what is the advantage of Friedman's H-statistic for some interaction compared to computing the functional decomposition for $\fh$?
    % Computes only the single component necessary, and may also give some (although bad) information in case the features are not independent.

\end{enumerate}