\textbf{Solution Quiz:}\\\noindent
\medskip

\begin{enumerate}
    \item What is the prediction target in contrast to the prediction?
    \begin{itemize}
    	\item[$\Rightarrow$] 
    \end{itemize}
    \item What is the model's mechanism in contrast to the data generating process (DGP)?
    \begin{itemize}
    	\item[$\Rightarrow$]  The DGP is the true underlying process generating the data. The model is estimated by optimizing the prediction performance, i.e. the model's mechanism tries to be a copy of the DGP but this mostly is not possible.
    \end{itemize}
    \item Yes or No: Shapley values and SHAP values are different names for the same concept.
    \begin{itemize}
    	\item[$\Rightarrow$] No % Begründung missing
    \end{itemize}
    \item What are SHAP value functions $v$ in contrast to SHAP values $\phi$?
    \begin{itemize}
    	\item[$\Rightarrow$] SHAP value function: a function assigning a value to a coalition
    	\item[$\Rightarrow$] SHAP values: Contribution of a feature to a prediction
    \end{itemize}
    \item What is the difference between marginal and conditional SHAP?
    \begin{itemize}
    	\item[$\Rightarrow$] The features that are not part of the coalition or the feature of interest are sampled from the marginal or the conditional distribution, respectively.
    \end{itemize}
    \item What is the difference between kernel SHAP and tree SHAP?
    \begin{itemize}
    	\item[$\Rightarrow$] 
    \end{itemize}
    \item Does the dependence structure in the DGP influence the SHAP result?
    \begin{itemize}
    	\item[$\Rightarrow$] 
    \end{itemize}
%    \item What is the motivation of LIME?
%    \begin{itemize}
%    	\item[$\Rightarrow$] Interpret a model locally by fitting a simple model (independent of the complexity of the global ML model). $\rightarrow$ local surrogate model
%    \end{itemize}
%    \item Both SHAP and LIME can be seen as local linear approximations. Name differences between SHAP and LIME.
%    \begin{itemize}
%    	\item[$\Rightarrow$] 
%    \end{itemize}
%    \item Name hyperparameters for LIME. What do they steer?
%    \begin{itemize}
%    	\item[$\Rightarrow$] 
%    \end{itemize}
\end{enumerate}
